A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2020-08-26 Papers/Amlap_2020_talk/amlap_2020_talk.tex
more work on talk

\documentclass[12pt, xcolor=beamer,table,usenames,dvipsnames, ignorenonframetext, ngerman]{beamer}
\usetheme{Frankfurt}
\usecolortheme{dove}
\usepackage{appendixnumberbeamer}
%\setbeamersize{text margin left=20pt,text margin right=20pt,}

@@

@@

\begin{frame}
\maketitle
%TODO maybe put a picture of Roger's face??, at least acknowledge him 
%Pre-amble -- thank Roger, outline with be methods heavy, and then talk about the specific experiment & how it's relevant to how we think about surprisal 
\end{frame}

% But first -- Incremental processing methods

\begin{frame}{Incremental processing methods}
\pause
Goal: Measure processing difficulty

Assumption: Longer RT = more difficulty
\medskip

\pause
\end{frame}

\begin{frame}{Incremental processing methods}
	How to measure RT?

Usual Methods
\begin{columns}[t]\pause
	\begin{column}{.5\textwidth}
		\begin{center}
			\textbf{\Large Eye-tracking}

@@

			\includegraphics[width=.9\textwidth]{../eye-tracker.jpeg}
			\pause
		\end{center}
	\end{column}\pause
	\begin{column}{.5\textwidth}
		\begin{center}
\textbf{\Large Self-paced reading}

@@

\begin{tikzpicture} %TODO on slide etc
\draw (-.5,-.5) rectangle (3.2,.5);
\onslide<4>{\node at (0,0) {The};}
\onslide<4>{\node at (1.7,0) {- - - - - - - - - - -};}
\onslide<5>{\node at (0,0) {- - - };}
\onslide<5>{\node at (.7,0) {cat};}
\onslide<5>{\node at (2,0) {- - - - - - - - };}
\onslide<6>{\node at (0.3,0) {- - - - - - };}
\onslide<6>{\node at (1.6,0) {drank};}
\onslide<6>{\node at (2.6,0) {- - - -};}
\onslide<7->{\node at (.9,0) {- - - - - - - - - - -};}
\onslide<7->{\node at (2.7,0) {milk.};}
\end{tikzpicture}
\end{center}
\onslide<8->{
}
\end{column}
\end{columns}
\bigskip
Different methods have different trade-offs
\end{frame}

\begin{frame}{An alternative: Maze}
%TODO First mentioned, then more elaborated in 2009 paper, has been used ~dozen times 
\centering
\tikzset{
	font={\fontsize{14pt}{12}\selectfont}}
%TODO Maybe update with circles to show the selections. 
\begin{tikzpicture}
\onslide<1-2>{\node  (c) at (-1,.5) {The};}
\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
\onslide<3-4>{\node  (d) at (1,.5) {dog};}
\onslide<5-6>{\node  (c) at (-1,.5) {revise};}
\onslide<5-6>{\node  (d) at (1,.5) {chased};}
\onslide<7-8>{\node  (c) at (-1,.5) {the};}
\onslide<7-8>{\node  (d) at (1,.5) {wish};}
\onslide<9-10>{\node  (c) at (-1,.5) {mitigate.};}
\onslide<9-10>{\node  (d) at (1,.5) {squirrel.};}
\onslide<2,8>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,align=center] at (c) {};}
\onslide<4,6,10>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,
align=center] at (d) {};}
\end{tikzpicture}
\end{frame}

\begin{frame}{An alternative: Maze}
\begin{columns}
	\begin{column}{0.5\textwidth}
		\begin{center}
		\textbf{\large G-maze}
		\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../gmaze.pdf}
		\end{center}
	\pause
	\end{column}
	\begin{column}{0.5\textwidth} 
		\begin{center}
		\textbf{\large L-maze}
			\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../lmaze.pdf}
		\end{center}
	\end{column}
\end{columns}

@@

Claim: forces incremental processing (no spillover)

\begin{flushright}
	{\small(Forster et al. 2009; Witzel et al. 2012)}
\end{flushright}
% Compelling results especially for G-maze, but 
% was run in the lab, and generating G-maze materials is a huge amount of work
\end{frame}

\begin{frame}{Using Maze}
Is Maze a viable alternative to web SPR?

Needs:
\begin{itemize}
	\item Run on web 
	\item Easy generation of distractor words
	\item Work for long or multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
\end{itemize} 
\pause

@@

\begin{frame}{Run on web}
	%TODO include screenshot
	Wrote an Ibex module
	\begin{itemize}
		\item Based off SPR module
		\item Fits in with existing workflow
	\end{itemize}
 Replicated Witzel et al. (2012) results (see Boyce et al. 2020)
\end{frame}
%
\begin{frame}{Using Maze}
	Is Maze a viable alternative to web SPR?

	Needs:
	\begin{itemize}
		\item Run on web 
		\item Easy generation of distractor words
		\item Work for long or multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
	\end{itemize} 
	\pause

\end{frame}
%
\begin{frame}{Generating distractors}
\pause
Task: Given  \textit{The dog chased} Find a word that can't be a continuation.
\begin{itemize}
	\item Tedious (and hard!) to do by hand. 
\end{itemize}

What makes something an unacceptable continuation?
\begin{itemize}
	\item Ungrammatical
	\item ...or otherwise really unlikely 
	\item $\approx$ high surprisal 
\end{itemize}

Can we use Neural Language Models?
\end{frame}
%
\begin{frame}{Can we use LMs?}
Language models (LMs)
\begin{itemize}
	\item Trained on large corpora to predict the next word
	\item Given a partial sentence, return probabilities of the next word
\end{itemize}

LSTM RNNs (ex. Gulordava 2018) show some syntactic knowledge %TODO cite people

Add some controls:
\begin{itemize}
\item Restrict to a list of possible distractors
\item Only consider length, frequency matches
\end{itemize}

\end{frame}
%
\begin{frame}{Can we use LMs to choose distractors?}

Does it work?

Yes, at least well enough.
\begin{itemize}
	\item Sometimes generates plausible distractors.
	\item Overall results comparable with G-maze %TODO citations to me, Sloggett
	%TODO show graph from Boyce et al
\end{itemize} 
\medskip

Parameter-tweaking may help, as would hand-checking. 
\end{frame}
%
\begin{frame}{Using Maze}
	Is Maze a viable alternative to web SPR?

	Needs:
	\begin{itemize}
		\item Run on web %TODO figure out how to make checkmarks again
		\item Easy generation of distractor words 
		\item Work for long or multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
	\end{itemize} 
	\pause

@@

Errors terminate sentences. 
Some error rate is inevitable.

Want to run multi-sentence items. 
\begin{itemize}
	\item Treat whole story as a unit: Almost no one makes it to the end.
	\item Treat each sentence as a unit:
	Many participants miss key context. 
\end{itemize}

@@

\end{frame}
%
\begin{frame}{Maze with Error Correction}
\centering
\tikzset{
	font={\fontsize{14pt}{12}\selectfont}}

@@

\begin{columns}
	\begin{column}{0.3\textwidth}
		\begin{center}
		\includegraphics[clip, trim=9cm 12.5cm 10cm 1cm,width=\textwidth]{../maze_diagram.pdf}
		\end{center}

	\end{column} \pause 
	\begin{column}{0.7\textwidth} 
		\begin{center}
			Features
			\begin{itemize}
				\item Easily implemented as a toggle in Ibex \pause
				\item Can look at long materials (in principle)
				\item Have all the data \pause
				\item Identify bad participants v bad distractors 
				%easy to weed out participants who aren't paying attention 
				\item Makes it less of a big deal when we have bad distractors
			\end{itemize}
		\end{center}
	\end{column}
\end{columns}
\medskip
\pause

Side effect: Can now run multi-sentence items!
\end{frame}
%
%%So now that I've explained the method, going to switch over to current experiment & what we hoped to get from it
%
\begin{frame}
	Various open questions we want to address -- really test suitability of Maze
	\begin{itemize}
		\item Do people comprehend/remember what they read?
		\item Does this error correction thing work? 
		\item Are people willing to read texts in this way? 
		\item Do we get predictability effects? (Both SPR, eye-tracking show certain surprisal effects, does Maze?)
	\end{itemize}
\end{frame}

\begin{frame}{Natural Stories}
Natural stories corpus (Futrell et al. 2017)
\begin{itemize}
	\item 10 stories, each about 1000 words
	\item Some unusual constructions, but read fluently
	\item 6 comprehension questions per story
\end{itemize}

@@

\begin{frame}{Participant accuracy}
100 participants from MTurk each read 1 story (20 minutes) \pause
\begin{center}
\includegraphics[width=.8\textwidth]{../error.pdf}
\end{center}
\end{frame}
\begin{frame}{Comprehension questions}
\begin{center} \pause
\includegraphics[width=.8\textwidth]{../comp.pdf}
\end{center}
% exclusion options: use comprehension checks as participant filter. Here we can use overall accuracy as the "paying attention" filter. Don't have to worry about how to write good comprehension questions. May still need to do RT exclusions for double-presses or getting distracted. 
\end{frame}

\begin{frame}{Surprisal Effects}
%TODO copy from updated abstract
Use 3 LMs to estimate surprisal: smoothed 5-gram, Gulordava RNN, Transformer-XL (Gulordava et al. 2018, Dai et al. 2019)
\pause

>>>>

\documentclass[12pt, xcolor=beamer,table,usenames,dvipsnames, ignorenonframetext, ngerman,t]{beamer}
\usetheme{Frankfurt}
\usecolortheme{dove}
\usepackage{appendixnumberbeamer}
%\setbeamersize{text margin left=20pt,text margin right=20pt,}

@@

\newcommand{\bluecheck}{}%
\DeclareRobustCommand{\greencheck}{%
	\tikz\fill[scale=0.6, color=ForestGreen]
	(0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;%
}

@@

\begin{frame}
\maketitle
%TODO maybe put a picture of Roger's face??, at least acknowledge him 
\end{frame}

\begin{frame}{Incremental processing methods}
\pause
Want a measure of processing difficulty \greencheck\pause

Assume that longer RT = more difficulty \pause
\smallskip
\begin{center}
{\large How to measure RT?}
\end{center}
\end{frame}
\begin{frame}{Incremental processing methods}

{\large Common ways to measure RT}
\begin{columns}[t]\pause
	\begin{column}{.5\textwidth}
		\begin{center}
			\textbf{\large Eye-tracking}

@@

			\includegraphics[width=.9\textwidth]{../eye-tracker.jpeg}
			\pause
		\end{center}
	\end{column}\pause
	\begin{column}{.5\textwidth}
		\begin{center}
\textbf{\large Self-paced reading}

@@

\begin{tikzpicture}
\draw (-.5,-.5) rectangle (3.2,.5);
\onslide<4>{\node at (0,0) {The};}
\onslide<4>{\node at (1.7,0) {- - - - - - - - - - -};}
\onslide<5>{\node at (0,0) {- - - };}
\onslide<5>{\node at (.7,0) {cat};}
\onslide<5>{\node at (2,0) {- - - - - - - - };}
\onslide<6>{\node at (0.3,0) {- - - - - - };}
\onslide<6>{\node at (1.6,0) {drank};}
\onslide<6>{\node at (2.6,0) {- - - -};}
\onslide<7->{\node at (.9,0) {- - - - - - - - - - -};}
\onslide<7->{\node at (2.7,0) {milk.};}
\end{tikzpicture}
\end{center}
\end{column}
\end{columns}
\bigskip
\onslide<8->{Different methods have different trade-offs}
\end{frame}

\begin{frame}[c]{An alternative: Maze}

\pause 
\centering

\bigskip

\tikzset{
	font={\fontsize{14pt}{12}\selectfont}}
\begin{tikzpicture}
\onslide<1-2>{\node  (c) at (-1,.5) {The};}
\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
\onslide<3-4>{\node  (d) at (1,.5) {dog};}
\onslide<5-6>{\node  (c) at (-1,.5) {revise};}
\onslide<5-6>{\node  (d) at (1,.5) {chased};}
\onslide<7-8>{\node  (c) at (-1,.5) {the};}
\onslide<7-8>{\node  (d) at (1,.5) {wish};}
\onslide<9-10>{\node  (c) at (-1,.5) {mitigate.};}
\onslide<9-10>{\node  (d) at (1,.5) {squirrel.};}
\onslide<2,8>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,align=center] at (c) {};}
\onslide<4,6,10>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,
align=center] at (d) {};}
\end{tikzpicture}
\end{frame}

\begin{frame}{An alternative: Maze}
\vskip -1em
{\small(Forster et al. 2009; Witzel et al. 2012)}
	\vskip -1em
\begin{columns}
	\begin{column}{0.5\textwidth}
		\begin{center}
		\textbf{\large G-maze}\\
		`Grammatical' choices\\
		\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../gmaze.pdf}
		\end{center}
	\pause
	\end{column}
	\begin{column}{0.5\textwidth} 
		\begin{center}
		\textbf{\large L-maze}\\
		`Lexical' choices\\
			\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../lmaze.pdf}
		\end{center}
	\end{column}
\end{columns}

@@

Claim: forces incremental processing (no spillover)
% Compelling results especially for G-maze, but 
% was run in the lab, and generating G-maze materials is a huge amount of work
\end{frame}

\begin{frame}{Maze Made Easy}

	Want to use Maze instead of web SPR\pause

	Needs some tweaks:\pause
	\begin{itemize}
		\item Run on web \pause
		\item Easily generate distractors \pause
		\item Work for multi-sentence items \pause
	\end{itemize}

@@

\begin{frame}{Run on web}
	Wrote an Ibex module \pause

	\begin{center}\includegraphics[width=.7\textwidth]{screenshot.png} \end{center}\pause

 Replicated Witzel et al. (2012) results (Boyce et al. 2020)
\end{frame}
%
\begin{frame}{Maze Made Easy}

	Want to use Maze instead of web SPR

	Needs some tweaks:
	\begin{itemize}
		\item Run on web \greencheck
		\item Easily generate distractors
		\item Work for multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
	\end{itemize}

\end{frame}
%
\begin{frame}{Generating distractors}
\pause
Goal: Find a word that can't continue a partial sentence
\begin{itemize}
	\item Ex.  \textit{The dog chased} \pause
	\item Tedious (and hard!) to do by hand
\end{itemize} \pause

What makes something an unacceptable continuation? \pause
\begin{itemize}
	\item Ungrammatical \pause
	\item ...or otherwise really unlikely \pause 
	\item $\approx$ high surprisal 
\end{itemize} \pause

Can we use Neural Language Models?
\end{frame}
%
\begin{frame}{Can we use LMs?}
	\pause
Language models (LMs)
\begin{itemize}
	\item Trained to predict the next word
	\item Given a partial sentence, return probabilities of the next word
\end{itemize}\pause

LSTM RNNs show some syntactic knowledge %TODO cite people 
\pause

Add some restrictions:
\begin{itemize}
\item Restrict to a list of possible distractors
\item Only consider length, frequency matches
\end{itemize}

\end{frame}
%
\begin{frame}{Does it work?} 
\pause
\vskip -2em
{\large Yes, at least well enough.} \pause
\begin{itemize}
	\item Sometimes generates plausible distractors. \pause
	\item A-maze results comparable with G-maze {\small(Boyce et al 2020, Sloggett et al 2020)} \pause 
\end{itemize} 
\medskip
\centering

\includegraphics[width=.7\textwidth]{../a_all.pdf}
\end{frame}
%
\begin{frame}{Maze Made Easy}

	Want to use Maze instead of web SPR

	Needs some tweaks:
	\begin{itemize}
		\item Run on web \greencheck
		\item Easily generate distractors \greencheck
		\item Work for multi-sentence items 
	\end{itemize}

@@

Want to run multi-sentence items. \pause

Problem: Errors terminate sentences. \pause
\begin{itemize}
	\item Treat whole story as a unit: ~ No one makes it to the end. \pause
	\item Treat each sentence as a unit: Some participants miss key context. \pause
\end{itemize}

@@

\end{frame}
%
\begin{frame}[c]{Maze with Error Correction}
\centering
\tikzset{
	font={\fontsize{14pt}{12}\selectfont}}

@@

\begin{columns}
	\begin{column}{0.4\textwidth}
		\begin{center}
		\includegraphics[clip, trim=9cm 12.5cm 10cm 1cm,width=\textwidth]{../maze_diagram.pdf}
		\end{center}

	\end{column} \pause 
	\begin{column}{0.6\textwidth} 
		\begin{center}
			\begin{itemize}
				\item Can be toggled in Ibex \pause
				\item Long materials feasible \pause
				\item Have all the data \pause
				\item Compensates for bad distractors
			\end{itemize}
		\end{center}
	\end{column}
\end{columns}
\medskip
\pause
\end{frame}
%

\begin{frame}{Maze Made Easy}

	Want to use Maze instead of web SPR

	Needs some tweaks:
	\begin{itemize}
		\item Run on web \greencheck
		\item Easily generate distractors \greencheck
		\item Work for multi-sentence items \greencheck ? 
	\end{itemize}

\end{frame}
%%So now that I've explained the method, going to switch over to current experiment & what we hoped to get from it
%
\begin{frame}{Current experiment}
	Various open questions to address \pause
	\begin{itemize}
		\item Are people willing to read long texts in this way? \pause
		\item Do people comprehend what they read? \pause
		\item Does the error correction variant work? \pause
		\item Do we get predictability effects? 
	\end{itemize}
\end{frame}

\begin{frame}{Natural Stories}
Natural stories corpus (Futrell et al. 2017) \pause
\begin{itemize}
	\item 10 stories, each about 1000 words \pause
	\item Some unusual constructions, but read fluently \pause
	\item 6 comprehension questions per story \pause
\end{itemize}

@@

\begin{frame}{Participant accuracy}
100 participants from MTurk each read 1 story (20 minutes) \pause
\begin{center}
\includegraphics[width=.8\textwidth]{../error.pdf}
\end{center}
\end{frame}

\begin{frame}{Comprehension questions}
\begin{center} \pause
\includegraphics[width=.8\textwidth]{../comp.pdf}
\end{center}
% exclusion options: use comprehension checks as participant filter. Here we can use overall accuracy as the "paying attention" filter. Don't have to worry about how to write good comprehension questions. May still need to do RT exclusions for double-presses or getting distracted. 
\end{frame}

%HERE%
\begin{frame}{Surprisal Effects}
%TODO copy from updated abstract
Use 3 LMs to estimate surprisal: smoothed 5-gram, Gulordava RNN, Transformer-XL (Gulordava et al. 2018, Dai et al. 2019)
\pause
